---
title: "Data Manipulation"
output: github_document
---

This file is for doing data manipulation.

```{r setup}
library(tidyverse)
```

load the litters and pups datasets.

```{r}
litters_df =
  read_csv(
    "data/FAS_litters.csv",
    na = c("", "NA", "."))

litters_df = janitor::clean_names(litters_df)

pups_df =
  read_csv(
    "data/FAS_pups.csv",
    na = c("", "NA", "."))

pups_df = janitor::clean_names(pups_df)
```

## `select` some variables

```{r}
select(litters_df, group, litter_number)

select(litters_df, group:gd_of_birth)

select(litters_df, -group)

select(litters_df, starts_with("gd"))

select(litters_df, starts_with("gd"),group)

select(litters_df, GROUP = group, everything())
```

a couple of alternatives

```{r}
rename(litters_df, GROUP = group)

relocate(litters_df, litter_number, gd0_weight)
```

In the pups data, select litter number, sex, and PD ears. 

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter` some rows

```{r}
filter(litters_df, gd_of_birth ==19)

filter(litters_df, pups_born_alive <=5)

filter(litters_df, pups_born_alive != 6)

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, group %in% c("Con7", "Con8"), pups_born_alive < 7)

filter(litters_df, pups_born_alive < 4 | pups_born_alive > 8)

drop_na(litters_df, gd0weight)
```

In the pups dataset, keep only rows where PD walk is less than 11 and sex is 2.
```{r}
filter(litters_df, pd_wwalk < 11, sex == 2)
```

## `mutate` to change some columns

let's add and change some columns

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)


mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

In the pups dataset, create a variable that subtracts 7 from PD walk

```{r}
mutate(pups_df, pd_walk_minus_7 = pd_walk - 7)
```

```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, desc(pups_born_alive))
```

## Do multiple steps...


THIS IS AN INCORRECT WAY TO WRITE THIS!!!! DON'T DO THIS!
```{r}
litters_df = read_csv("data/FAS_litters.csv", na = c("","NA","."))
litters_rename_df = janitor::clean_names(litters_df)
litters_with_gd_df = select(litters_rename_df, group, starts_with("gd"))
litters_no_na_df = drop_na(litters_with_gd_df)
litters_wt_gain_df = mutate(litters_no_na_df, wt_gain = gd18_weight - gd0_weight)
```

Nesting is worse.

Correct way:
```{r}
litters_df =
  read_csv("data/FAS_litters.csv", na = c("","NA",".")) |> 
  janitor::clean_names() |> 
  select(group, starts_with("gd")) |> 
  drop_na() |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
  )

```

Load pups, clean names, drop missing, keep litter number and pd variables, add pd walk - 7

```{r}
pups_df =
  read_csv("data/FAS_pups.csv", skip = 3, na = c("","NA",".")) |> 
  janitor::clean_names() |> 
  drop_na() |> 
  select(litter_number, starts_with("pd")) |> 
  mutate(
    pd_walk_minus_7 = pd_walk - 7
  )
  
```













